---
title: "Web Scraping with `rvest`"
subtitle: "Math 241, Week 5"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#Load web scraping library
library(rvest)
library(tidyverse)

```

# Grab Tables from the Web

Let's grab the **Portland area sports teams** table on Portland's [Wikipedia page](https://en.wikipedia.org/wiki/Portland,_Oregon).

```{r}
#Store url
url <- "https://en.wikipedia.org/wiki/Portland,_Oregon"

## Scrape html and store table

#Grab all the tables and then navigate to the one you wanted.
tables <- url %>%
  read_html() %>%
  html_nodes(css = "table")

#Grab the specific table
champ_table <- html_table(tables[[8]], fill = TRUE)
champ_table

```

## Population data over time

Using the same approach, grab the table of Portland's population over time (table 4), and create a simple visualization of the total population count over time.


# Another Example

* Although we saw that we can use `datapasta` to grab these data, let's scrape the [NYTimes.com's College Access Index](https://www.nytimes.com/interactive/2017/05/25/sunday-review/opinion-pell-table.html) table.

```{r}
# Store url
url <- "https://www.nytimes.com/interactive/2017/05/25/sunday-review/opinion-pell-table.html"

## Scrape html and store table

# Grab the table
tables <- url %>%
  read_html() %>%
  html_nodes(css = "table")

#Grab the specific table
college_access_table <- html_table(tables[[1]], fill = TRUE)

#Option 2: Use the specific css
college_access_table2 <- url %>%
  read_html() %>%
  html_node(css = "#opinion-pell-table > div > div.g-item.g-sortable-table > table") %>%
  html_table()


```

## Create a visualization

Choose 3 variables and create a heuristic visualization using this dataset.

```{r}

#making variable names useable
collacc.clean <- college_access_table %>% 
  rename("freshman.class" = "Freshman class",
         "pell.grad.share" = "Pell grad share",
         "netprice.mid" = "Net price, mid. income",
         "college.access.index" = "College Access Index",
         "endowment" = "Endowment per student") 


collacc.clean.fix <- collacc.clean 
#removing commas from freshman.class in order to convert to numeric
collacc.clean.fix$freshman.class <- gsub(',', '', collacc.clean.fix$freshman.class)
#removing $ and k from netprice.mid in order to convert to numeric
collacc.clean.fix$netprice.mid <- gsub('\\$|k', '', collacc.clean.fix$netprice.mid)
#converting to numeric
collacc.clean.fix <- collacc.clean.fix %>% 
  mutate(freshman.class = as.numeric(freshman.class),
         netprice.mid = as.numeric(netprice.mid) * 1000)

# Function to convert strings like "$11.1k" to numeric
convert_to_numeric <- function(x) {
  # Remove '$'
  x <- gsub('\\$', '', x)
  # Convert to numeric
  if (grepl('k', x)) {
    as.numeric(gsub('k', '', x)) * 1000
  } else if (grepl('m', x)) {
    as.numeric(gsub('m', '', x)) * 1000000
  } else {
    as.numeric(x)
  }
}

# Apply the function to the endowment column
collacc.clean.fix$endowment <- sapply(collacc.clean.fix$endowment, convert_to_numeric)




ggplot(data = collacc.clean.fix, aes(x = freshman.class, y = netprice.mid)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

